---
title: "gamma generalized linear models"
description: "(gone bayesian)"
date: "11/10/22"
image: ""
---

What makes GLMs "general" is that

1.  The response can be modeled as coming from any distribution in the *exponential family*
2.  A link function that connects the linear component of our model to the response distribution. Sometimes it's helpful for me to think of the link function as making the linear component and the response distribution "compatible". (ex: if our model outcomes are in (0,1) then we need a link function that will map that restricted space onto the entire real line).

In the case of just a normal linear model it's pretty obvious what th

So maybe we are fitting a model to a response variable that's continuous, strictly positive, and right skewed.

```{r, echo=F, message=F, warning = F}
library(tidyverse)
data <- data.frame(
  x = rgamma(1000, 1.5, 10)
)
```

```{r}
data %>% 
  ggplot(aes(x = x)) +
  geom_density(fill = "lightblue", alpha = 0.4) +
  theme_bw()
```

Seems like this data structure would merit the use of a gamma glm- the response certainly would be well described by a gamma distirbution (in fact the plot above actually was generated with random draws from a `Gamma(1.5, 10)` random variable) and so we might 